---
title: "resios"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(resios)
```

# RESIOS

To retrieve data you need to [ask for it via email](consultasios@ree.es).
You can find the official [help page for the API](https://api.esios.ree.es/doc/index.html) with many examples.

Once you get the token I recommend to save it in your .Renviron file (if you only will use R to retrieve data)

```{r esios-archives}
get_token()
ea <- esios_archives()
head(ea)
```

Besides the archives you can get daily data from the more than 1900 indicators available:

```{r esios-indicators, eval=FALSE}
# Will fail due to esios error on their server.
ei_df <- esios_search_indicators()
dim(ei_df)
head(ei_df)
```

Some of them are really interesting, like the price per MWh

```{r pvpc}
ei <- esios_pvpc()
library("ggplot2")
ggplot(ei) +
  geom_path(aes(tz_time, value, col = geo_name, linetype = geo_name)) +
  theme_minimal() +
  labs(y = "€/MWh", x = "Hour", title = "Price per time", 
       col = "Region", linetype = "Region") +
  scale_y_continuous(sec.axis = sec_axis(trans = function(x){x/1000}, name = "€/KWh", 
                                         labels = scales::label_number(accuracy = 0.01, suffix = "€"))) +
  scale_x_datetime(date_breaks = "4 hours", date_labels = "%H") +
  theme(legend.position = "bottom", legend.direction = "horizontal")
  
```

# REE

You can also query the Red Electrica Española which have also very interesting data.

```{r}
rc0 <- ree_call(widget = "demanda-tiempo-real", category = "demanda")
plot(rc0$`Demanda real`$datetime, rc0$`Demanda real`$value, type = "l", pch = 2,
     main = "Electricity demand in peninsular Spain", 
     xlab = "Time", ylab = "Value (?)")
lines(rc0$`Demanda programada`$datetime, rc0$`Demanda programada`$value, 
      type = "l", col = "blue", lty = 2)
lines(rc0$`Demanda prevista`$datetime, rc0$`Demanda prevista`$value, 
      type = "l", col = "red", lty = 3)
abline(h = Sys.time(), col = "black")
legend("topleft", legend = c("real", "foreseen", "programmed"), 
       fill = c("black", "red", "blue"))
```

